\LARGE \bf Learning Depth-Sensitive Conditional Random Fields for Semantic Segmentation of RGB-D Images

### 2013-09-14 paper.tex
Bildunterschrift, rote Kringel.
git-svn-id: https://svn.ais.uni-bonn.de/svn/staff/publications/icra_2014_mueller_nyu@4375 b2511e17-9ac0-488b-bc21-1d6d82e5baa2
redo related work, criple citations.
git-svn-id: https://svn.ais.uni-bonn.de/svn/staff/publications/icra_2014_mueller_nyu@4378 b2511e17-9ac0-488b-bc21-1d6d82e5baa2

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
    We present a structured learning approach to semantic annotation of RGB-D
    images. 
    Our method learns to reason about spatial relations of objects and
    fuses low-level class predictions to a consistent interpretation of a
    scene. Our model incorporates color, depth and 3D scene features, on which
    an energy function is learned to directly optimize object class prediction using
    the loss-based maximum-margin principle of structural support vector machines.
    %With the proliferation of cheap % TODO urghs formulation
    %high-quality RGB-D sensors, it became possible to capture more and more
    %data not only as color images, but with additional depth information. The
    %depth channel provides important clues not only for scene structure, but
    %also for predicting semantic object classes.
    %We demonstrate how conditional random field learning, an approach
    %that has shown great promise for RGB images, can exploit depth information
    %to improve semantic interpretation of the scene.
    We evaluate our approach on the NYU V2 dataset of indoor scenes, a challenging
    dataset covering a wide variety of scene layouts and object classes. We
    hard-code much less information about the scene layout into our model then previous
    approaches, and instead learn object relations directly from the data. We
    find that our conditional random field approach improves upon previous
    work, setting a new state-of-the-art for the dataset.
\end{abstract}

@@

We thereby extend the success of learned conditional random field models for
semantic segmentation in RGB images to the domain of 3D scenes.
Our emphasis lies on exploiting the additional depth and 3D information in all
processing steps, while relying on \emph{learning} to create a model that 
is adjusted to the properties of the sensor input and environment.

@@

\begin{figure}
    \begin{center}
        \includegraphics[width=\linewidth]{images/teaser}
    \end{center}
    \caption{%
        Visualization of the method proposed in this paper see \Secref{learning} for details.\label{teaser}
    \figlabel{height}}
\end{figure}

\section{RELATED WORK}
The task of dense semantic labeling in natural images has been widely studied
in computer vision.  Our approach of using pixel-based low-level cues is
similar to the work of \citet{shotton2006textonboost}, who introduced
TextonBoost for semantic segmentation, and used a hand-coded CRF energy to
model spatial interactions. More recent work usually employ learned conditional
random fields, learning object interactions from the
data to directly maximize a loss function~\citep{lucchi2013learning, krahenbuhl2012efficient},
which is the approach we are following.

Work semantic segmentation and mapping using RGB-D data only recently emerged with
the availability of the NYU Depth Dataset V1~\citep{silberman2011indoor}.
\citet{silberman2011indoor} captured a wide variety of indoor scenes using a structured light sensor,
and provided dense semantic object class annotation. Their work also introduced a baseline method
for semantic segmentation of RGB-D image, which used an approach based on superpixels, interest
point descriptors and a conditional random field. While pairwise potentials for the CRF were carefully
designed for the dataset, potentials were either directly set by hand or estimated using empirical
frequencies. This is in contrast to our work, which applies structured prediction techniques
to learn potentials automatically to optimize predictive performance.
\citet{ren2012rgb} focus on the design of features for semantic labeling of RGB-D data, and use
a hierarchical segmentation to provide context. While they also define a conditional random field on superpixels,
their model is again not learned, but a weighted Potts model, using only a
probability of boundary map, and not taking spatial layout into account at all.

\citet{SilbermanECCV12} extended the NYU Depth Dataset V1 to the NYU Depth Dataset V2 that we are also
using in this work.  Their focus is on inferring support relations in indoor
scenes, such as objects resting on tables or shelfs, which in turn rest on the floor.
Their approach is based on robust estimation of 3D plane hypotheses, which are then jointly
optimized with support relations and structure classes.
\citet{SilbermanECCV12} use a complex pipeline, employing significant domain knowledge.
In this work, on the other hand, we try to learn all relevant domain specific features
directly from the data, which allows us to out-perform the work of \citet{SilbermanECCV12}
with respect to structure class segmentation.

@@

\citet{stueckler2013} proposed a real-time system for online learning and prediction of semantic classes.
Their main contribution is the online integration of low-level predictions over multiple frames
by estimating camera motion and building a 3D model of the scene.
We use the implementation of random forest provided by \citet{stueckler2013}, but instead of
integrating predictions over time, we focus on exploiting the structure within a single frame.

\section{LEARNING DEPTH-SENSITVE CONDITIONAL RANDOM FIELDS}\seclabel{learning}
We take a conditional random field approach, whose nodes represent a labeling of
superpixels.  We use an energy consisting of first and second order factors
(also called unary and pairwise potentials), with learned potential functions.
Let us denote the representation of an input image by $x$ and a labeling of
superpixels into semantic classes by $y$.  Then the general form of the energy is
\begin{equation}\eqlabel{energy}
    g(x, y) = \sum_{v \in V} \psi_v(x, y_v) + \sum_{(v, w) \in E} \psi_{v,w}(x, y_{v}, y_{w}).
\end{equation}
Here $V$ enumerates the superpixels, and $E\subset V \times V$ is a set of
edges, encoding adjacence between superpixels.

We learn the unary and pairwise energy functions $\psi_v$ and $\psi_{v, w}$ from the training data
using a structural support vector machine~\citep{joachims2009cutting}. 
The concept of structural support vector machines allows for a principled,
maximum-margin based, loss-sensitive training of conditional random fields. Learning the potential yields much
more complex interactions than the simple Potts potentials that are often used in the literature.

In general, structural support vector machines learn the parameters $\theta$ of a predictor of the form
\begin{equation}\eqlabel{prediction}
f(x) = \argmax_{y \in \mathcal{Y}} \theta^T \Phi(x, y).
\end{equation}
We choose $\psi$ in \Eqref{energy} to be linear in the learnable parameters and
the data-depended features, resulting in a form equivalent to
\Eqref{prediction}. Our features are described in detail below.  We use the
$1$-slack formulation of the structural SVM~\citep{joachims2009cutting} and
solve the maximization in \Eqref{prediction} using a combination of fusion
moves~\citep{lempitsky2010fusion} and the AD$^3$ algorithm of
\citet{martins2011augmented}.  In contrast to graph-cut inference, fusion moves
and AD$^3$ can work with arbitrary potential function, and allow precise
learning using the structural support vector machine approach.

@@

All hyper parameters were adjusted using 5-fold cross-validation. The
hyper parameters of the random forests were found using the \texttt{hyperopt}
framework of \cite{bergstra2011algorithms}. For the CRF model, the only
hyper-parameters are related to the superpixel segmentation, and the single
hyper-parameter $C$ of the structural SVM formulation. These were adjusted
using standard grid-search.
We found 500 superpixels per image to work best, which allow for a maximum
possible performance of 95\% average class accuracy on the validation set.

\subsection{Results}\seclabel{results}
\Tabref{results} compares different components of our approach with the literature.
Please note that we \emph{first} designed our final model, using only the
validation data. We now report accuracies of simpler models for reference,
however these results were not used for model selection. To separate
the influence of loss-based training and the spatial reasoning of the conditional random field,
we also train a standard support vector machine on the unary potentials for comparison.

The random forest prediction, as reported in \citet{stueckler2013} is already quite competitive.
Grouping into superpixels slightly improves performance, by removing
high-frequency noise and snapping to object boundaries. Somewhat surprisingly,
using a standard unstructured SVM with rescaled loss already advances the mean
accuracy decidedly above the previous state-of-the art. We
attribute this mostly to the ability of the SVM to exploit correlation between
classes and uncertainty within the superpixels.
Additionally, the SVM has access to the ``height'' feature, that was not
included in the random forest.
This performance is still improved upon, both in class average and pixel
average performance by the learned conditional random field approach, yielding
the best published result so far for both measures. The increase over the
standard SVM is $1.5\%$ for class average accuracy and $2.0\%$ for pixel
average accuracy.

@@

We found that our approach improves results most for scenes with a clear geometric structure,
which is not surprising. We see that evidence from the random forest is often very noisy, and biased
away from the ``props'' class. While the unstructured SVM can correct somewhat
for the class imbalance, it has no way to make larger areas consistent, which
the conditional random field can.
On the other hand, performance of the CRF deteriorates slightly on very crowded
scenes with a mixture of small furniture and prop objects, as
can be seen in the two right-most images. In these scenes, depth information is
often noisy, and it is hard to make geometric statements on the superpixel
level. As the input from the random forest is also often of low quality for
crowded scenes, the conditional random field has little chance to recover.

@@

    \includegraphics[width=\linewidth]{images/00845_gt.png}&%
    \includegraphics[width=\linewidth]{images/00781_gt.png}&%
    \includegraphics[width=\linewidth]{images/01331_gt.png}&%
    \includegraphics[width=\linewidth]{images/00118_gt.png}&%
    \includegraphics[width=\linewidth]{images/01203_gt.png}\\
    \multicolumn{5}{@{}c@{}}{Ground Truth}\vspace{3mm}\\
    \multicolumn{5}{@{}c@{}}{
        \includegraphics[]{legend.pdf}
    }
    \end{tabu}
\caption{%
Qualitative evaluation of the conditional random field.
The first three images illustrate errors in the original prediction that can be corrected, while
the second two images illustrate failure modes. Pixels marked as void
are excluded from the evaluation. See the \secref{results} for details.
\figlabel{comparison}}
\end{figure*}

\section{SUMMARY AND DISCUSSION}
We introduce a conditional random field formulation for semantic segmentation
of structure classes in RGB-D images. We base our model in the output of an efficient GPU implementation
of random forest, and model spatial neighborhood using a superpixel-based approach.
We combine color, depth and 3D orientation features into an energy function that is fully learned using 
 the structural support vector machine approach.  By explicitly modeling 3D
 relations in a fully learned framework, we improve the state-of-the art on the
 NYU V2 dataset for semantic annotation of structure classes.

While our approach allows modeling of spatial relations, these are limited to
local interactions. In future work, these interactions could be extended to
larger areas using latent variable models~\citep{dann2012pottics} or higher
order potentials~\citep{kohli2009robust}.
Another possible line of future work is to combine our approach with a more
task-specific one, directly including support plane assumptions into the model,
as done by \citet{SilbermanECCV12}. Finally, we could also combine our
single-frame approach with the approach of \citet{stueckler2013}, which fuses
individual views in 3D to exploit temporal coherence.

While this work did not explicitly address real time application, the random
forest implementation that we build upon allows for real-time
processing~\citep{stueckler2013}. The SLIC superpixel algorithm can also be
implemented on GPU in real-time, as was demonstrated by \citep{ren2011gslic}, and similarly
the normal features we use also have
real-time capabilities~\citep{holz_2011_robocup}.
Finally, fusion move inference for our model is very efficient for our model,
opening up the possibility to implement our approach entirely in real time.

\printbibliography{}

\end{document}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
    We present a structured learning approach to semantic annotation of RGB-D
    images. 
    Our method learns to reason about spatial relations of objects and
    fuses low-level class predictions to a consistent interpretation of a
    scene. Our model incorporates color, depth and 3D scene features, on which
    an energy function is learned to directly optimize object class prediction using
    the loss-based maximum-margin principle of structural support vector machines.
    %With the proliferation of cheap % TODO urghs formulation
    %high-quality RGB-D sensors, it became possible to capture more and more
    %data not only as color images, but with additional depth information. The
    %depth channel provides important clues not only for scene structure, but
    %also for predicting semantic object classes.
    %We demonstrate how CRF learning, an approach
    %that has shown great promise for RGB images, can exploit depth information
    %to improve semantic interpretation of the scene.
    We evaluate our approach on the NYU V2 dataset of indoor scenes, a challenging
    dataset covering a wide variety of scene layouts and object classes. We
    hard-code much less information about the scene layout into our model then previous
    approaches, and instead learn object relations directly from the data. We
    find that our conditional random field approach improves upon previous
    work, setting a new state-of-the-art for the dataset.
\end{abstract}

@@

We thereby extend the success of learned CRF models for
semantic segmentation in RGB images to the domain of 3D scenes.
Our emphasis lies on exploiting the additional depth and 3D information in all
processing steps, while relying on \emph{learning} to create a model that 
is adjusted to the properties of the sensor input and environment.

@@

\begin{figure}
    \begin{center}
        \includegraphics[width=\linewidth]{images/teaser}
    \end{center}
    \caption{%
        Overview of the proposed semantic segmentation method.\label{teaser}
    \figlabel{height}}
\end{figure}

\section{RELATED WORK}
The task of dense semantic annotation of 3D maps has seen an increased interest
in recent years.  Early work includes \citet{nuchter2008towards}, who combine
6D SLAM, surface annotation and object recognition, to build semantically
annotated maps. They demonstrate their approach on a mobile robot in an indoor
environment.
%
More recently \citet{sengupta2012automatic} introduced a dataset of
semantically annotated street-scenes in a closed track, captured as pairs of
stereo images. They approach the task by jointly reasoning about 3D layout and
semantic of the scenes and produce a dense labeling on image level.  In
\citet{senguptaurban}, they extend their approach to produce a volumetric
reconstruction of the scene, together with a dense semantic labeling of the
volumetric representation. Their approach to image segmentation is build on
the hierarchical CRF apporach of
\citet{ladicky2009associative}, which is similar in spirit to our approach,
but uses Potts potentials together with cross-validation to set potentials.

Recent approaches for indoor semantic annotation of maps mostly focused on
RGB-D images, which are now easy to obtain using structured light sensors.
\citet{stuckler2012semantic} used a Random Forest model to obtain a dense
semantic labeling of images, integrating predictions over multiple views in 3D.
They evaluated their approach on a limited set of table-top and indoors scenes.
%
\citep{silberman2011indoor} introduced the NYU Depth Dataset V1 indoor dataset,
which consisted of a large variety of densely annotated indoor scences, captured as RGB-D images.
Their work also introduced a baseline method for semantic segmentation of RGB-D
image, which used an approach based on superpixels, interest point descriptors
and a CRF.
While pairwise potentials for the CRF were carefully
designed for the dataset, potentials were either directly set by hand or estimated using empirical
frequencies. This is in contrast to our work, which applies structured prediction techniques
to learn potentials automatically to optimize predictive performance.
%
\citet{ren2012rgb} evaluated the design of features for semantic labeling of RGB-D data, and use
a hierarchical segmentation to provide context. While they also define a CRF on superpixels,
their model is again not learned, but a weighted Potts model, using only a
probability of boundary map, and not taking spatial layout into account at all.
%
\citet{SilbermanECCV12} extended the NYU Depth Dataset V1 to the NYU Depth Dataset V2 that we are also
using in this work.  Their focus is on inferring support relations in indoor
scenes, such as objects resting on tables or shelfs, which in turn rest on the floor.
Their approach is based on robust estimation of 3D plane hypotheses, which are then jointly
optimized with support relations and structure classes.
\citet{SilbermanECCV12} use a complex pipeline, employing significant domain knowledge.
In this work, on the other hand, we try to learn all relevant domain specific features
directly from the data, which allows us to out-perform the work of \citet{SilbermanECCV12}
with respect to structure class segmentation.

@@

\citet{stueckler2013} extends the approach of \citet{stuckler2012semantic} to
a real-time system for online learning and prediction of semantic classes. Their
use a GPU implementation of random forests, and integrate 3D scene information
in an online fashion. They evaluate their approach on the dataset of
\citet{SilbermanECCV12} with promising results.
We use the implementation of random forest provided by
\citet{stueckler2013}, but instead of integrating predictions over time, we
focus on exploiting the structure within a single frame.

While many of the works mentioned in this chapter make use of a CRF
approach, we are not aware of any prior work on semantic annotation of 3D maps
that fully learns their potentials. This allows our approach to infer
the scene structure from the data, without the need to specify relevant relations
by hand.
%The task of dense semantic labeling in natural images has been widely studied
%in computer vision.  Our approach of using pixel-based low-level cues is
%similar to the work of \citet{shotton2006textonboost}, who introduced
%TextonBoost for semantic segmentation, and used a hand-coded CRF energy to
%model spatial interactions. More recent work usually employ learned conditional
%random fields, learning object interactions from the
%data to directly maximize a loss function~\citep{lucchi2013learning, krahenbuhl2012efficient},
%which is the approach we are following.

\section{LEARNING DEPTH-SENSITVE CONDITIONAL RANDOM FIELDS}\seclabel{learning}
We take a CRF approach, whose nodes represent a labeling of
superpixels.  We use an energy consisting of first and second order factors
(also called unary and pairwise potentials), with learned potential functions.
Let us denote the representation of an input image by $x$ and a labeling of
superpixels into semantic classes by $y$.  Then the general form of the energy is
\begin{equation}\eqlabel{energy}
    g(x, y) = \sum_{v \in V} \psi_v(x, y_v) + \sum_{(v, w) \in E} \psi_{v,w}(x, y_{v}, y_{w}).
\end{equation}
Here $V$ enumerates the superpixels, and $E\subset V \times V$ is a set of
edges, encoding adjacence between superpixels.

We learn the unary and pairwise energy functions $\psi_v$ and $\psi_{v, w}$ from the training data
using a structural support vector machine (SSVM)~\citep{joachims2009cutting}. 
The concept of SSVMs allows for a principled,
maximum-margin based, loss-sensitive training of CRFs. Learning the potential yields much
more complex interactions than the simple Potts potentials that are often used in the literature.

In general, structural SSVMs learn the parameters $\theta$ of a predictor of the form
\begin{equation}\eqlabel{prediction}
f(x) = \argmax_{y \in \mathcal{Y}} \theta^T \Phi(x, y).
\end{equation}
We choose $\psi$ in \Eqref{energy} to be linear in the learnable parameters and
the data-depended features, resulting in a form equivalent to
\Eqref{prediction}. Our features are described in detail below.  We use the
$1$-slack formulation of the structural SVM~\citep{joachims2009cutting} and
solve the maximization in \Eqref{prediction} using a combination of fusion
moves~\citep{lempitsky2010fusion} and the AD$^3$ algorithm of
\citet{martins2011augmented}.  In contrast to graph-cut inference, fusion moves
and AD$^3$ can work with arbitrary potential function, and allow precise
learning using the SSVM approach.

@@

All hyper parameters were adjusted using 5-fold cross-validation. The
hyper parameters of the random forests were found using the \texttt{hyperopt}
framework of \cite{bergstra2011algorithms}. For the CRF model, the only
hyper-parameters are related to the superpixel segmentation, and the single
hyper-parameter $C$ of the structural SVM formulation. These were adjusted
using grid-search.
We found 500 superpixels per image to work best, which allow for a maximum
possible performance of 95\% average class accuracy on the validation set.

\subsection{Results}\seclabel{results}
\Tabref{results} compares different components of our approach with the literature.
Please note that we \emph{first} designed our final model, using only the
validation data. We now report accuracies of simpler models for reference,
however these results were not used for model selection. To separate
the influence of loss-based training and the spatial reasoning of the CRF,
we also train a usual support vector machine (SVM) on the unary potentials for comparison.

The random forest prediction, as reported in \citet{stueckler2013} is already quite competitive.
Grouping into superpixels slightly improves performance, by removing
high-frequency noise and snapping to object boundaries. Somewhat surprisingly,
using a standard unstructured SVM with rescaled loss already advances the mean
accuracy decidedly above the previous state-of-the art. We
attribute this mostly to the ability of the SVM to exploit correlation between
classes and uncertainty within the superpixels.
Additionally, the SVM has access to the ``height'' feature, that was not
included in the random forest.
This performance is still improved upon, both in class average and pixel
average performance by the learned CRF approach, yielding
the best published result so far for both measures. The increase over the
standard SVM is $1.5\%$ for class average accuracy and $2.0\%$ for pixel
average accuracy.

@@

We found that our approach improves results most for scenes with a clear geometric structure,
which is not surprising. We see that evidence from the random forest is often very noisy, and biased
away from the ``props'' class. While the unstructured SVM can correct somewhat
for the class imbalance, it has no way to make larger areas consistent, which
the CRF can.
On the other hand, performance of the CRF deteriorates slightly on very crowded
scenes with a mixture of small furniture and prop objects, as
can be seen in the two right-most images. In these scenes, depth information is
often noisy, and it is hard to make geometric statements on the superpixel
level. As the input from the random forest is also often of low quality for
crowded scenes, the CRF has little chance to recover.

@@

    \includegraphics[width=\linewidth]{images/00845_gt.png}&%
    \includegraphics[width=\linewidth]{images/00781_gt.png}&%
    \includegraphics[width=\linewidth]{images/01331_gt.png}&%
    \includegraphics[width=\linewidth]{images/00118_gt.png}&%
    \includegraphics[width=\linewidth]{images/01203_gt.png}\\
    \multicolumn{5}{@{}c@{}}{Ground Truth}\vspace{3mm}\\
    \multicolumn{5}{@{}c@{}}{
        \includegraphics[]{legend.pdf}
    }
    \end{tabu}
\caption{%
Qualitative evaluation of the CRF.
The first three images illustrate errors in the original prediction that can be corrected, while
the second two images illustrate failure modes. Pixels marked as void
are excluded from the evaluation. See the \secref{results} for details.
\figlabel{comparison}}
\end{figure*}

\section{SUMMARY AND DISCUSSION}
We introduce a CRF formulation for semantic segmentation
of structure classes in RGB-D images. We base our model in the output of an efficient GPU implementation
of random forest, and model spatial neighborhood using a superpixel-based approach.
We combine color, depth and 3D orientation features into an energy function that is fully learned using 
 the SSVM approach. By explicitly modeling 3D
 relations in a fully learned framework, we improve the state-of-the art on the
 NYU V2 dataset for semantic annotation of structure classes.

While our approach allows modeling of spatial relations, these are limited to
local interactions. In future work, these interactions could be extended to
larger areas using latent variable models or higher
order potentials~\citep{ladicky2009associative}.
Another possible line of future work is to combine our approach with a more
task-specific one, directly including support plane assumptions into the model,
as done by \citet{SilbermanECCV12}. Finally, we could also combine our
single-frame approach with the approach of \citet{stueckler2013}, which fuses
individual views in 3D to exploit temporal coherence.

While this work did not explicitly address real time application, the random
forest implementation that we build upon allows for real-time
processing~\citep{stueckler2013}. The SLIC superpixel algorithm can also be
implemented on GPU in real-time, as was demonstrated by \citep{ren2011gslic}, and similarly
the normal features we use also have
real-time capabilities~\citep{holz_2011_robocup}.
Finally, fusion move inference for our model is very efficient for our model,
opening up the possibility to implement our approach entirely in real time.
\printbibliography{}
\end{document}
